---
title: "DATA MANIPULATION IN R"
author: "BASIT et JANET"
date: "3/30/2022"
output:
  html_document:
    ---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATASET: OrchardSprays
## MANIPULATING DATA


OrchardSprays {datasets}	R Documentation
Potency of Orchard Sprays
Description
An experiment was conducted to assess the potency of various constituents of orchard sprays in repelling honeybees, using a Latin square design.

A data frame with 64 observations on 4 variables.

[,1]	rowpos	numeric	Row of the design
[,2]	colpos	numeric	Column of the design
[,3]	treatment	factor	Treatment level
[,4]	decrease	numeric	Response
Details
Individual cells of dry comb were filled with measured amounts of lime sulphur emulsion in sucrose solution. Seven different concentrations of lime sulphur ranging from a concentration of 1/100 to 1/1,562,500 in successive factors of 1/5 were used as well as a solution containing no lime sulphur.

The responses for the different solutions were obtained by releasing 100 bees into the chamber for two hours, and then measuring the decrease in volume of the solutions in the various cells.

An 8 x 8 Latin square design was used and the treatments were coded as follows:

A	highest level of lime sulphur
B	next highest level of lime sulphur
.	
.	
.	
G	lowest level of lime sulphur
H	no lime sulphur

###Data Manipulation on OrchardSprays Dataset
```{r OrchardSprays}
library(dplyr)
library(knitr)
data("OrchardSprays")
head(OrchardSprays)
attach(OrchardSprays)
```


```{r echo=TRUE}
###### Filter Treatment level
filt1 = filter(OrchardSprays, treatment == "A")
print(filt1)
```
```{r}
filt2 = filter(OrchardSprays, treatment == "B")
print(filt2)
```
```{r}
### Filter Response greater than 100
filt3 = filter(OrchardSprays, decrease > 100)
print(filt3)
```
```{r}
#### Filter Response less than 100
filt4 = filter(OrchardSprays, decrease < 100)
head(filt4)
```

```{r}
#### Filter Row of the design
filt5 = filter(OrchardSprays, rowpos == 2)
print(filt5)
```

```{r}
#### Filter Column of the design
filt6 = filter(OrchardSprays, colpos == 4)
print(filt6)
```

```{r}
#### Arrange in Descending Order
Arr = arrange(OrchardSprays, desc(decrease))
head(Arr)
```

```{r}
#### Select treatment and decrease from OrchardSprays
TrD = select(OrchardSprays, treatment, decrease)
head(TrD)
```

```{r}
#### Mutation
Sum = mutate(OrchardSprays, total = sum(decrease))
head(Sum)
```
```{r}
Aver = mutate(OrchardSprays, mean = sum(decrease)/64)
head(Aver)
```

```{r}
Orc = OrchardSprays %>%
  mutate(ratio = (decrease/sum(decrease)),
         Perc = ratio*100)
head(Orc)
```
```{r}
##### Summarization
Orch = OrchardSprays %>%
  select(decrease,treatment)%>%
  filter(decrease>50)%>%
  summarise(ave = mean(decrease))
print(Orch)
```

```{r, echo=TRUE}
#### The percentage of soluton obtained by releasing 100 bees into the chamber for 2hours
grouping = OrchardSprays %>%
  group_by(decrease) %>%
  summarise(total = sum(decrease),
            AVE = mean(decrease),
            perc = decrease [64]/sum(decrease, na.rm = TRUE)*100)
print(grouping)
```


###DATA VISUALIZATION WITH GGPLOT2 PACKAGE
####Making a boxplot of data
#####HISTOGRAM

```{r HISTOGRAM, echo=FALSE}
library(ggplot2)
ggplot(data=OrchardSprays,aes(x=decrease))+
  geom_freqpoly()+
  theme_grey()+
  labs(title="OrchardSprays", subtitle="Decrease", x="30")
  hist(decrease,col="cyan")
```

##BOXPLOT
```{r}
Spray=ggplot(OrchardSprays,aes(factor(rowpos),decrease))+
geom_boxplot()+
ylab("decrease") + 
xlab("rowpos")
Spray

####BOXPLOT
boxplot(decrease~rowpos+colpos,data=OrchardSprays)
```

###Running Anova
```{r}
mydata=data.frame(decrease,rowpos, colpos, treatment)
mydata
myfit=lm(decrease~ rowpos+treatment+colpos, mydata)
anova(myfit)
a=plot(decrease~ rowpos+treatment+colpos, mydata)

a
```

###BAR CHART
```{r}
ggplot(data=OrchardSprays,aes(x=decrease))+
  geom_bar()+
  theme_classic()

###BAR PLOT
barplot(decrease~rowpos+colpos,data=OrchardSprays)
```





